Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
finish models

$$B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\text{ where } a,b>0.$$
We extend the equation to **non-negative real number $c_i$** by applying a linear approximation

@@

>>>>

$$B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\text{ where } a,b>0.$$

We extend the equation to **non-negative real number $c_i$** by applying a linear approximation

@@

.center[
# 💪Power of Visual Tests💪
]

The power of a visual test may depend on the **ability of the particular subject** (Majumder, Hofmann, and Cook 2013).

- We assume the **individual skill has negligible effect** on the variation of the power to **simplify the model structure**, thereby **obviate a costly large-scale experiment** to estimate complex covariance matrices.

We use the **logistic regression** to estimate the power:

$$Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),$$

where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.

- The **effect size $E$** is the only predictor.
- The intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.

.center[
# 🧪Experimental Design🧪
]

An experiment is conducted to investigate the difference between **conventional hypothesis testing** and **visual testing** in the application of **linear regression diagnostics**.

We focus on two types of residual departures - **Non-linearity** and **Heteroskedasticity**.

Overall, we collected **7974 evaluations** on **1152 unique lineups** performed by **443 subjects** throughout three data collection periods.

.center[
# 🧪Experimental Design🧪
]

**Non-linearity model**:

$$\begin{aligned} \boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\ \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\ \boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\ \boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)), \end{aligned}$$
where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are length $n$ vectors, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as

$$g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \text{ for }  k > 0. $$

The null regression model used to fit the realizations generated by the above model is formulated as

$$\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},$$

where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

.center[
# 🧪Experimental Design🧪
]

**Heteroskedasticity model**:

$$\begin{aligned} \label{eq:heter-model} \boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\ \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\ \boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), \end{aligned}$$

where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{\varepsilon}$ are length $n$ vectors and $g(.)$ is the scaling function defined in the previous slide.

The null regression model is also provided in the previous slide.

```{r}
magick::image_read_pdf("images/different-shape-of-herimite-1.pdf", pages = 1)
magick::image_read_pdf("images/different-sigma-1.pdf", pages = 1)
```

.center[
.caption[
Fig. 5: _Non-linearity forms generated for the residual plots are controlled by the parameter \\(j\\). And the strength of the signal of the non-linearity model is controlled by \\(\sigma\\)._
]
]

.image-center[

```{r out.width = "70%", fig.height = 2.67, fig.width = 8}
set.seed(10086)

stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

a_labels <- c("left-triangle (a = -1)", "butterfly (a = 0)", "right-triangle (a = 1)")

# Generate data for a = -1
dat_a_n1 <- heter_model(a = -1, 
                        x = {
                          raw_x <- rand_uniform(-1, 1);
                          closed_form(~stand_dist(raw_x))
                          },
                        b = 128)$gen(300) %>%
  mutate(a = a_labels[1])

# Generate data for other a
map(c(0, 1), function(a) {
  heter_model(a = a,
              x = {
                raw_x <- rand_uniform(-1, 1); 
                closed_form(~stand_dist(raw_x))
                }, 
              b = 128)$
    gen(300, computed = select(dat_a_n1, x, e)) %>%
    mutate(a = a_labels[a + 2])
}) %>%

  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(a = factor(a, levels = a_labels)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~a, scales = "free", ncol = 3) +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r}
magick::image_read_pdf("images/different-b-1.pdf", pages = 1)
```

.center[
.caption[
Fig. 5: _Heteroskedasticity forms generated for the residual plots are controlled by the parameter \\(a\\). And the strength of the signal of the heteroskedasticity model is controlled by \\(b\\)._
]
]
